CA Affordabulity

### 2020-09-17 paper.Rmd
update

```{r rates pop,echo=FALSE, include=TRUE, message=FALSE,warning=FALSE, results='asis'}
rates <- read_csv("Data/affordability.csv")
priv <- select(b,PWSID,Ownership,D_POPULATI)
rates <- left_join(rates,b,by=c("pwsid"="PWSID")) 
rates <- rates[which(!is.na(rates$Ownership)),]
ggplot(rates, aes(x=log10(D_POPULATI), fill=Ownership)) +
  geom_density(alpha=0.4)
# Add mean lines
ear13 <- read_tsv("Data/EAR/earsurveyresults_2013ry.zip",col_types="ccdcccdddc")
ear14 <- read_tsv("Data/EAR/earsurveyresults_2014ry.zip",col_types="ccdcccdddc")
ear15 <- read_tsv("Data/EAR/earsurveyresults_2015ry.zip",col_types="ccdcccdddc")
ear16 <- read_tsv("Data/EAR/earsurveyresults_2016ry.zip",col_types="ccdcccdddc")
ear17 <- read_tsv("Data/EAR/earsurveyresults_2017ry.zip",col_types="ccdcccdddc")
ear18 <- read_tsv("Data/EAR/earsurveyresults_2018ry.zip",col_types="ccdcccdddc")

ear <- bind_rows(ear13,ear14,ear15,ear16,ear17,ear18)
rm(ear13,ear14,ear15,ear16,ear17,ear18)
save(ear,file="Data/EAR/ear.rds")
ear$QCat<-substr(ear$QuestionName,1,2)

@@

## Shut-offs
```{r shutoffs}
conn <- ear %>% filter(SectionName=="Connections" & Year==2018)
conn_sfr <- conn %>% filter(QuestionName=="SFR Potable Total" & Year==2018)
conn_mfr <- conn %>% filter(QuestionName=="MFR Potable Total" & Year==2018)
conn_total <- conn %>% filter(QuestionName=="T Potable Total" & Year==2018) %>%rename(connections=QuestionResults) %>% mutate(connections=as.numeric(connections)) %>% select(PWSID,connections) %>% distinct(.keep_all=TRUE)

@@

shutoff_comments <- shutoffs %>% filter(QuestionName=="WR SHUT_OFFS Comments")
  re
shutoff_pa<- ear%>%filter(QuestionName=="WR SHUT_OFFS Payment Assistance Total" & Year==2018) %>%distinct(PWSID, QuestionResults) %>% rename(shutoff_pa=QuestionResults) %>% filter(!is.na(shutoff_pa)) %>% mutate(shutoff_pa=as.numeric(shutoff_pa))

@@

r_shutoff <- left_join(r,conn_shutoff,by="PWSID") %>% distinct(PWSID,connections,.keep_all=TRUE) %>%
ggplot(filter(r_shutoff), aes(x=size_cat, y=shutoff_rate, colour=Ownership)) +
  geom_boxplot()

lm()

@@

## Subsidy/ Cost assistance programs

@@

>>>>

```{r rates pop,echo=FALSE, include=TRUE, message=FALSE,warning=FALSE, results='asis'}
rates <- read_csv("Data/affordability.csv")
priv <- select(b,PWSID,Ownership,D_POPULATI)
rates <- left_join(rates,b,by=c("pwsid"="PWSID")) 
rates <- rates[which(!is.na(rates$Ownership)),]
ggplot(rates, aes(x=log10(D_POPULATI), fill=Ownership)) +
  geom_density(alpha=0.4)
# Add mean lines
# ear13 <- read_tsv("Data/EAR/earsurveyresults_2013ry.zip",col_types="ccdcccdddc")
# ear14 <- read_tsv("Data/EAR/earsurveyresults_2014ry.zip",col_types="ccdcccdddc")
# ear15 <- read_tsv("Data/EAR/earsurveyresults_2015ry.zip",col_types="ccdcccdddc")
# ear16 <- read_tsv("Data/EAR/earsurveyresults_2016ry.zip",col_types="ccdcccdddc")
# ear17 <- read_tsv("Data/EAR/earsurveyresults_2017ry.zip",col_types="ccdcccdddc")
# ear18 <- read_tsv("Data/EAR/earsurveyresults_2018ry.zip",col_types="ccdcccdddc")
# 
# ear <- bind_rows(ear13,ear14,ear15,ear16,ear17,ear18)
# rm(ear13,ear14,ear15,ear16,ear17,ear18)
# save(ear,file="Data/EAR/ear.rds")
load("Data/EAR/ear.rds")
ear$QCat<-substr(ear$QuestionName,1,2)

@@

## Shut-offs
```{r shutoffs, echo=FALSE, include=TRUE, message=FALSE,warning=FALSE, results='asis'}
conn <- ear %>% filter(SectionName=="Connections" & Year==2018)
conn_sfr <- conn %>% filter(QuestionName=="SFR Potable Total" & Year==2018)
conn_mfr <- conn %>% filter(QuestionName=="MFR Potable Total" & Year==2018)
conn_total <- conn %>% filter(QuestionName=="T Potable Total" & Year==2018) %>% rename(connections=QuestionResults) %>% mutate(connections=as.numeric(connections)) %>% select(PWSID,connections) %>% distinct(.keep_all=TRUE)

@@

shutoff_comments <- shutoffs %>% filter(QuestionName=="WR SHUT_OFFS Comments")

shutoff_pa<- ear%>%filter(QuestionName=="WR SHUT_OFFS Payment Assistance Total" & Year==2018) %>%distinct(PWSID, QuestionResults) %>% rename(shutoff_pa=QuestionResults) %>% filter(!is.na(shutoff_pa)) %>% mutate(shutoff_pa=as.numeric(shutoff_pa))

@@

r_shutoff <- left_join(r,conn_shutoff,by="PWSID") %>% distinct(PWSID,connections,.keep_all=TRUE) 
ggplot(r_shutoff, aes(x=size_cat, y=shutoff_rate, colour=Ownership)) +
  geom_boxplot()

r_shutoff_model <- lm(
  shutoff_rate ~ Private + log(D_POPULATI),
  data = filter(r_shutoff, D_POPULATI>3300)
#  family = quasibinomial
)

summary(r_shutoff_model)
```

## Subsidy/ Cost assistance programs
```{r subsidies}
subs <- ear %>% filter(substr(QuestionName,1,5)=="WR LL" & Year==2018)
subsidies <- ear %>% 
  filter(QuestionName=="WR LL Subsidies" & Year==2018) %>% 
  #filter(QuestionResults!="-") %>% 
  select(PWSID,QuestionResults) %>% 
  rename(subsidy=QuestionResults) %>% distinct(.keep_all=TRUE)

subsidies_LI <- subs %>% 
  filter(QuestionName=="WR LL Eligibility-Low Income" & Year==2018) %>% 
  filter(QuestionResults!="-") %>%
  select(PWSID,QuestionResults) %>%
  rename(LowIncome=QuestionResults) %>% distinct(.keep_all=TRUE) %>% mutate(LowIncome="Yes")

subsidies_D <- subs %>% 
  filter(QuestionName=="WR LL Eligibility-Disabled" & Year==2018) %>% 
  filter(QuestionResults!="-") %>%
  select(PWSID,QuestionResults) %>%
  rename(Disabled=QuestionResults) %>% distinct(.keep_all=TRUE)  %>% mutate(Disabled="Yes")

subsidies_S <- subs %>% 
  filter(QuestionName=="WR LL Eligibility-Seniors" & Year==2018) %>% 
  filter(QuestionResults!="-") %>%
  select(PWSID,QuestionResults) %>%
  rename(Seniors=QuestionResults) %>% distinct(.keep_all=TRUE) %>% mutate(Seniors="Yes")

subsidies_M <- subs %>% 
  filter(QuestionName=="WR LL Eligibility-Medical" & Year==2018) %>% 
  filter(QuestionResults!="-") %>%
  select(PWSID,QuestionResults) %>%
  rename(Medical=QuestionResults) %>% distinct(.keep_all=TRUE) %>% mutate(Medical="Yes")

subsidies_O <- subs %>% 
  filter(QuestionName=="WR LL Eligibility-Other" & Year==2018) %>% 
  filter(QuestionResults!="-") %>%
  select(PWSID,QuestionResults) %>%
  rename(Other=QuestionResults) %>% distinct(.keep_all=TRUE) %>% mutate(Other="Yes")

subsidies_HowFunded <- subs %>% 
  filter(QuestionName=="WR LL Funded") %>% 
  filter(QuestionResults!="-") %>%
  select(PWSID,QuestionResults) %>%
  rename(Disabled=QuestionResults) %>% distinct(.keep_all=TRUE)

subsidies_BenefitForm <- subs %>% 
  filter(QuestionName=="WR LL Benefit Form") %>% 
  filter(QuestionResults!="-") %>%
  select(PWSID,QuestionResults) %>%
  rename(Disabled=QuestionResults) %>% distinct(.keep_all=TRUE)

shutoff_pa<-left_join(b,shutoff_pa,by="PWSID")

subs_b <- b %>% 
  left_join(subsidies,by="PWSID") %>%
  left_join(subsidies_LI,by="PWSID") %>%
  left_join(subsidies_S,by="PWSID") %>%
  left_join(subsidies_D,by="PWSID") %>%
  left_join(subsidies_M,by="PWSID") %>%
  left_join(subsidies_O,by="PWSID")

subs_b$subsidy[which(subs_b$LowIncome=="Yes" | subs_b$Disabled=="Yes" | subs_b$Other=="Yes" | subs_b$Medical=="Yes" | subs_b$Seniors=="Yes")]<-"Yes"
subs_b$LowIncome[which(subs_b$subsidy=="No")]<-"No"
subs_b$Disabled[which(subs_b$subsidy=="No")]<-"No"
subs_b$Medical[which(subs_b$subsidy=="No")]<-"No"
subs_b$Seniors[which(subs_b$subsidy=="No")]<-"No"
subs_b$Other[which(subs_b$subsidy=="No")]<-"No"

subs_b$LowIncome[which(subs_b$subsidy=="Yes" & is.na(subs_b$LowIncome))]<-"No"
subs_b$Disabled[which(subs_b$subsidy=="Yes" & is.na(subs_b$Disabled))]<-"No"
subs_b$Medical[which(subs_b$subsidy=="Yes" & is.na(subs_b$Medical))]<-"No"
subs_b$Seniors[which(subs_b$subsidy=="Yes" & is.na(subs_b$Seniors))]<-"No"
subs_b$Other[which(subs_b$subsidy=="Yes" & is.na(subs_b$Other))]<-"No"

@@

@@

sd<-lm(serious_violator_last_decade ~ Ownership + log(D_POPULATI),
  data=filter(bv,D_POPULATI>3300)
)
summary(sd)
